Will Kamala Harris will win the 2024 Presidential election

### 2024-11-04 paper/paper.qmd
intro done

---
title: "Will Kamala Harris will win the 2024 Presidential election"
subtitle: "Analysing the polls using baysian modeling"
author: 
  - Yan Mezhiborsky
  - Cristina Burca
  - Shreya Sakura Noskor
thanks: "Code and data are available at: [https://github.com/Mezhi18/US_Election2024.git](https://github.com/Mezhi18/US_Election2024.git)."
date: today
date-format: long
abstract: "We analyse the probability of Kamala Harris winning in the 7 swing states of USA using baysian modeling; Arizona, Michigan, Pennsylvania, Nevada, Georgia, Wisconsin, North Carolina. USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of USA will help understand the future economic condiotion of the world. Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris. Provided this we predict that Kamal Harris will win the election. By creating a baysian model based on \"polls of polls \" where we compare the results from different polls, we are able to make this prediction. "
format:
  pdf:
    bibliography: references.bib
    documentclass: article
    geometry: margin = 1in
number-sections: true
toc: true
bibliography: references.bib
---

@@

In this paper we aim to analyze polling data related to the 2024 election, with a particular emphasis on the swing states. By examining various polls and identifying patterns in voter sentiment, we seek to uncover the factors that may influence voter behavior in these highly contested regions. Our analysis will explore demographic shifts, the impact of key issues, and the level of voter engagement across different swing states. Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the forces shaping the 2024 presidential election.

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@@

Next to look at the actual contents of the survey. They can be accessed through this link:[Sample survey question](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEL7G89G56QlHp_qqIPzGbJdUME9UQzJaWUlCVldZTktRV09UQ0Q3N0hQRi4u). There are 3 things that we were careful of when we were creating this survey. First thing we considered was transperency. People need the reassurence that the data that is being collected will not be used against them, and so we teel them what data we do collect and how the data we collect cannot be used to identify a person. The second thing we focus on is readability of the questions. We tried to make them as simple as possible using accessibile language, and tried to keep it short as well (@Tourangeau_Rips_Rasinski_2000). We also prioritised the size of the survey. We kept it to a short 11 questions that will tell us their political standing in the past and present. We know what current issues are important to them as well as their age group and the state they are from. This will help us gather data without inconveniencing the person.

Finally we take a look at how we sample from the gathered data. We thought about random sampling but we were worried about having unequal proportions of the demographic we would advertise to. So if more people that go to facebook responded, that doesn't necessairly mean that people use facebook more than the other platforms. So we we decided to combine startified and cluster  sampling(@annurev). The idea is that we will choose based on the different platforms we advertise to first to have a cluster sampling method. And then we will sample the data by dividing the people by the certain aspects of the surevy, like if they are in different age groups or if they are registered voters. This is called Stratified Sampling (@annurev). Then we take a union of all the samples and we gather a group that we are able to build models based off of.

>>>>

---
title: "Will Kamala Harris will win the 2024 Presidential election"
subtitle: "Analysing the polls using baysian modeling"
author: 
  - Yan Mezhiborsky
  - Cristina Burca
  - Shreya Sakura Noskor
thanks: "Code and data are available at: [https://github.com/Mezhi18/US_Election2024.git](https://github.com/Mezhi18/US_Election2024.git)."
date: today
date-format: long
abstract: "We analyse the probability of Kamala Harris winning in the 7 swing states of USA using baysian modeling; Arizona, Michigan, Pennsylvania, Nevada, Georgia, Wisconsin, North Carolina. USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of USA will help understand the future economic condiotion of the world. Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Wisconsin(48.38% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris. Provided this we predict that Kamal Harris will win the election. By creating a baysian model based on \"polls of polls \" where we compare the results from different polls, we are able to make this prediction. "
format:
  pdf:
    bibliography: references.bib
    documentclass: article
    geometry: margin = 1in
number-sections: true
toc: true
bibliography: references.bib
---

@@

In this paper we aim to analyze polling data related to the 2024 election, with a particular emphasis on the swing states. By examining various polls and identifying patterns in voter sentiment, we seek to uncover the factors that may influence voter behavior in these highly contested regions. Our analysis will explore demographic shifts, the impact of key issues, and the level of voter engagement across different swing states. Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the forces shaping the 2024 presidential election. We use Bayesian regression model to create a model to predict the percent of population that will support in the swing states based on pollster, the number of days since Kamala Harris joined the race. More on the model and the variables, in the Data and Model sections.

We noticed in most predictions of the presidential candidate they are usually based on one survey or pollster and they try to gather data from the one poll to predict the the future President. However, through looking at the results from different polls we are able to get a better, more accurate prediction. We also look at over state polls which allows us to focus on the swing states. Swing states are states that change sides every election. The 2 major parties that fight for power are the Republicans and the Democrats. Some states are always voting for one party regardless of the candidate so, the swing states are the once that tell us more about who might win the election. The gap we are trying to cover is predicting the future President of USA by looking at different polls, while being as effecient as possible by looking only at the swing states.

To go a little more in-depth into what we actually did, we have to first discuss our estimand. Our estimand, the parameter we aim to study, is the predicted support for Kamala Harris in the swing states (measured in %). We do this by creating Baysian model with 3 parameters; state, pollster, and says since Kamala Harris was announced as candidate. We have the priors set with a normal distribution centered at 0, with a standard deviation of 2.5. We expect that because the swing states are states that don't have a party that they dominantly vote for, the expected support will be around 50%. We also make other graphs in the Data section, which helps us understand the values themselves.

From the result section we find that the average predicted support for Kamala Harris for the 7 swing states were as follows: Arizona at 46.26%, Nevada at 46.43%, Georgia at 46.59%, North Carolina at 47.26%, Michigan at 47.30%, Pennsylvania at 48.02%, Wisconsin at 48.38%. Based on this we predict that Arizona, and Georgia will vote Donald trump and the rest of the states will vote Kamala Harris.

Predicting who will become president is important because it helps people understand and prepare for future changes in the country’s policies and priorities. A new president can shape laws, affect the economy, influence international relationships, and address key issues like healthcare, education, and climate change. Accurate predictions allow citizens, businesses, and governments to plan and adapt. For example, if polls indicate a particular candidate might win, groups can start thinking about how their plans might align with the candidate's proposed policies. It also helps keep people informed and engaged in the political process, making them more likely to vote and participate in democracy.
Results paragraph

The paper is structured with the introduction in the beginning followed by the Data section wich takes a look at the methodology of the paper and how the data was gathered. We explain the variables we isolated in order for this analysis to be complete. Then it is followed by the Model section where we introduce the model and explain the graphs of related models. The result sections will cover the model summary and summary table of the predictions for each of the swing states. The last section is the discussion section where we talk about what the results means and we give our final conclusion and weaknesses. We also give a final word about who will win the upcoming election. We also have additional appendices, where we talk about the methodology of the polls and how they compare. The second appendix talks about a survey we designed with reasons and disadvantage. The third appendix just shows the models in more depth.

@@

Next to look at the actual contents of the survey. They can be accessed through this link: [Sample survey question](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEL7G89G56QlHp_qqIPzGbJdUME9UQzJaWUlCVldZTktRV09UQ0Q3N0hQRi4u). There are 3 things that we were careful of when we were creating this survey. First thing we considered was transperency. People need the reassurance that the data that is being collected will not be used against them, and so we feel them what data we do collect and how the data we collect cannot be used to identify a person. The second thing we focus on is readability of the questions. We tried to make them as simple as possible using accessibile language, and tried to keep it short as well (@Tourangeau_Rips_Rasinski_2000). We also prioritised the size of the survey. We kept it to a short 11 questions that will tell us their political standing in the past and present. We know what current issues are important to them as well as their age group and the state they are from. This will help us gather data without inconveniencing the person. There are disadvanatages to the way we structured the survey. For example, we were keeping time and security in mind, so we restricted the survey to 11 questions, however, this limits the amount of answers we get from the reader. We may be missing some crucial information that helps us figure out which group of people may vote for whom, such as ethnicity and wages (@annurev).

Finally we take a look at how we sample from the gathered data. We thought about random sampling but we were worried about having unequal proportions of the demographic we would advertise to. So if more people that go to facebook responded, that doesn't necessarily mean that people use facebook more than the other platforms. So we we decided to combine stratified and cluster  sampling(@annurev). The idea is that we will choose based on the different platforms we advertise to first to have a cluster sampling method. And then we will sample the data by dividing the people by the certain aspects of the surevy, like if they are in different age groups or if they are registered voters. This is called Stratified Sampling (@annurev). Then we take a union of all the samples and we gather a group that we are able to build models based off of.
There are obvious disadvantages as well. For example, we have fewer questions meaning when we sample based on the question we might have cases where, some groups are over or under represented. This can happen for example when most of the people who answered might be from age 18-50 because those are the age groups with access to a phone and are frequently visiting the sites we advertise our survey on. Also the different apps may skew the data as Facebook has a older audience as opposed to Instagram and Tiktok. Another disadvantage is the fact that we have 2 different sampling methods that we use and we combine them both, meaning that this can be an expensive process. Doing so may exceed the allocated budget however, it will be worth it because we minimize the sampling bias we talk about earlier where we may have a group that has more access to the survey. The last disadvantage we also foresee is the fact that younger audience may not be as interested in filling out a survey about politics as they feel, it might not directly affect them. We try to minimize this by offering compensation for their time when they fill out the survey, i.e. free membership services for 3 months on Spotify.

Although there are both advatages and disadvantages, we feel that this survey is well though out and doesn't allow for ambiguity in the answers of the participants.
